---
title: "Cheng_Lab2_Session2"
author: "Lily Cheng"
date: "10/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressMessages(library(tidyverse))
library(ggpubr)
```

two conditions for z distribution: population standard distribution, population normal distribution (n>30, normally have normal distribution, can use t distribution.)


Part2. Proababilities for the Z-distribution

*pnorm*: Calculates the probability associated with an input value for a random variable

*qnorm*: Calculates the value of the random associated with an input probability

We're told the following:

The population distribution for cat weights (pounds) is normally distributed with a mean of 9.2 pounds and a sd of 1.4 pounds.

What is the probability of randomly selecting a cat from that population that:

- weighs less than 8 pounds?
- weights more than 10 pounds?
- weights between 9 and 12 pounds?

```{r}
cat_mean <- 9.2 # mean cat population weight (pounds)
cat_sd <- 1.4 # SD population weight (pounds)

# probability of a cat <8 pounds?
p_less_8 <- pnorm(8, mean = cat_mean, sd = cat_sd)

# the probability of randomly selecting a cat from this population that weighs less than 8 pounds is 0.196

# Probability of a cat > 10 pounds?
p_norm_10 < pnorm(10, mean = cat_mean, sd = cat_sd, lower.tail = FALSE)

# probability of randomly selecting a cat weighing more than 10 pounds is 0.284.

# What is the probability of randomly selecting a cat that weighs between 9 and 12 pounds?

p_norm_9_12 <- pnorm(12, mean = cat_mean, sd = cat_sd) - pnorm(9, mean = cat_mean, sd = cat_sd)

# There is a 0.53 probability of randomly selecting a cat weighing between 9 and 12 pounds from this population.


# What is the probability of randomly selecting a cat that weighs more than 13 pounds?

p_norm_13 <- pnorm(13, mean = cat_mean, sd = cat_sd, lower.tail = FALSE)

```

The probability of randomly selecting a cat that weighs less than 8 pounds is `r {p_less_8}`


How can we find values of a random variable associated with a specified percentile?

Use *qnorm*

At what weight would a cat be at the 80th percentile in weight for the population?

```{r}
cat_80 <- qnorm(0.8, mean = cat_mean, sd = cat_sd)
# a cat weighing 10.4 pounds is at the 80th percentile in weight for population.

# Between the 40th and 60th percentile, what are the endpoints at that range?

cat_60 <- qnorm(0.6, mean = cat_mean, sd = cat_sd) 
cat_40 <- qnorm(0.4, mean = cat_mean, sd = cat_sd)
```

## part3. T-distribution for confidence intervals and hypothesis testing

We'll be using the *t.test* function for all of it.

using dataset "iris" built in R based package.

### a. Go exploring (histogram and qq-plots with facet wrap)

```{r}
iris_hist <- ggplot(iris, aes(x = Petal.Length)) +
  geom_histogram(aes(fill = Species))+
  facet_wrap(~ Species, scales = "free") + # if there are two variables we are intested in e.g.     facet_wrap(State ~ Species)
  theme_classic()

iris_hist

# geom_qq, aes(sample = )

iris_qq <- ggplot(iris, aes(sample = Petal.Length))+
  geom_qq(aes(color = Species))+
  facet_wrap(~ Species)
iris_qq
```

Create a dataset of the "iris" datasets that only includes the setosa irises, only keep the column for petal length, rename that column to "length"

```{r}
setosa_petals <- iris %>%
  filter(Species == "setosa") %>% 
  select(Petal.Length) %>% 
  rename(length = Petal.Length)

mean_length <- mean(setosa_petals$length)
mean_length # 1.462 cm

```

```{r}
iris_ci <- t.test(setosa_petals$length)
iris_ci

```
I see a claim that the mean length of setosa iris petals is 1.5 cm. I want to test claim using Edgar Anderson's sample

Two-sided:

H0: Mean = 1.5 cm
HA: Mean is not 2.5 cm

```{r}
iris_t_1 <- t.test(setosa_petals$length, mu = 1.5)
iris_t_1
```

Claim: Mean length is 1.5 cm.
You think: Mean length is less than 1.5 cm.

One-sided test:
H0: Mean is not less than 1.5 cm

```{r}

t_test_oneway <- t.test(setosa_petals$length, mu = 1.5, alternative = "less") # alternative define a one tail test, then the confidence interval does not make sense, because it just goes infinite to one direction. It has to be tail to have confidence interval
t_test_oneway
```

## Part4. Two-Sample T-test

```{r}
flint <- read.cs("")

flint_t <- t.test(flint$pb_1, flint_t$pb_3, paired = TRUE)
flint_t

# lead concentrations after initial sampling (mean = ??) after 2min of flushing water lines (near = ??) differ significantly (t(270) = 6.37, p < 0.001)
```

