---
title: "lab5_Cheng"
author: "Lily Cheng"
date: "10/31/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(pwr)
library(knitr)
library(kableExtra)
library(plotly)
library(extrafont)
library(ggrepel)
```


First we'll do an F-test for equal variances to see if we can overfide the default Welch approximation.
```{r}
# F-test for equal variances
# 
chickwts_f <- chickwts %>% 
  filter(feed == "horsebean" | feed == "linseed") %>% 
  var.test(weight ~ feed, data = .)

chickwts_f

# Retain the null hypothesis variances are equal
# override the default in t.test (var.equal = FALSE)
# t.test()

chickwts_t <- chickwts %>% 
  filter(feed == "horsebean" | feed == "linseed") %>% 
  t.test(weight ~ feed, data = ., var.equal = TRUE)

chickwts_t

## OR:

#chick_simple <- chickwts %>% 
 # var.test(weight ~ feed, data = .)  # something wrong with the lines

#t_test <- chick_simple %>% 
 # t.test(weight ~ feed, data = ., var.equal = TRUE)

#t_test

chick_simple <- chickwts %>% 
  filter(feed == "horsebean" | feed == "linseed")

f_test <- chick_simple %>% 
  var.test(weight ~ feed, data = .)

t_test <- chick_simple %>% 
  t.test(weight ~ feed, data = .)

t_test
```
p value means, if there is no difference between the two population, that the probability of the two mean are different is p.

Mean weights of chickens fed horsebean (mean sd) and linseed (mean sd) differ significantly (t(`r chickwts_t$parameter`) = `r
round(chickwts_t$statistic,2)`, *p* = `r round(chickwts_t$p.value, 3)`, $\alpha$=0.05).

### Part2. Power analysis

We need to collect samples to test a hypothesis (planning on doing a 2-sample t-test) to see if there is a difference in means for phosphate concentrations in lagoons downstream from golf courses of NOT downstream from golf courses. 

* A priori* power analysis: give you an estimate of how many observations you'll need to collect (really useful for justifying budgets, etc. for proposals)

- n: sample size
- power: power
- sig.level: alpha
_ d: effect size (Cohen's *d*)

Find n for "small" effect size (~0.2), and a "large" effect size (~0.8) as endpoints for n estimations

```{r}
# A priori power analysis

# How many samples needed if there is a SMALL effect size, and alpha = 0.05. power = 0.80

power_small <- pwr.t.test(n = NULL, power = 0.8, sig.level = 0.05, d = 0.2)
power_small

power_large <- pwr.t.test(n = NULL, power = 0.8, sig.level = 0.05, d = 0.8)
power_large

```
###Part3. Piping into ggplot

```{r}
mortality <- read_csv("drug_mortality.csv")
income <- read_csv("state_income.csv")
```

```{r}
mortality_1 <- mortality %>% 
  filter(ages == "All Ages",
         sex == "Both Sexes",
         race_and_hispanic_origin == "All Races-All Origins",
         state != "United States") %>% 
  select(year, state:death_rate)

m15 <- mortality_1 %>% 
  filter(year == 2015) %>% 
  mutate(highlight = ifelse(state == "Kentucky", "Yes", "No")) %>% 
  arrange(-death_rate) %>% 
  head(10) %>% 
  ggplot(aes(x = reorder(state, -death_rate), y = death_rate))+
  geom_col(aes(fill = highlight))+
  scale_fill_manual(values = c("gray60", "red"))

m15

ggplotly(m15)

# graph_1
```

###4. Interactive bubble plot

We'll use full_join (dyplyr) - this means that you retain ALL observations, even if the level only exists in one of the data frames

```{r}
View(mortality_1)
mi_2015 <- full_join(mortality_1, income) %>% 
  filter(year == 2015)

graph_3 <-  ggplot(filter(mi_2015, med_income > 60000), 
         aes(x = med_income,
             y = death_rate,
             label = state))+
  geom_point()+
  geom_text_repel()

graph_3 <-  ggplot(filter(mi_2015), 
         aes(x = med_income,
             y = death_rate,
             label = state))+
  geom_point(aes(size = population, color = state), alpha = 0.4)+
  geom_text_repel(size = 3)+
  theme(legend.position = "NA")

graph_3

ggplotly(graph_3)
```

### Part5. table using kable/kableExtra

```{r}
ca_table <- mortality_1 %>% 
  filter(year >= 2010,
         state == "California")

ca_table

ca_final <- kable(ca_table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

ca_final
```

