---
title: "Lab4"
author: "Lily Cheng"
date: "10/24/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressMessages(library(tidyverse))
library(effsize)
library(pwr)
library(knitr)
```

```{r}
nc_births <- read_csv("nc_birthweights.csv")
View(nc_births)
```

```{r}
nc_new <- nc_births %>% 
  mutate(
    sdp = case_when(
      smoke == 0 ~"nonsmoker",
      smoke == 1 ~ "smoker"
    )
  ) %>% 
  filter(sdp != "NA") %>% 
  select(sdp, gained, weeks, tpounds)
```

###3. Bubbleplot in ggplot

A bubbleplot is just a scatterplot graph where the size of the points depend on the value of a selected variable.
three steps in making plots
1 tell it we are using ggplot 2 tell it where we put the data 3 tell it what type of plot we want it to make

```{r}
bw_scatter <- ggplot(nc_new, aes(x = weeks, y = tpounds)) +
  geom_point(aes(color = sdp, size = gained), alpha = 0.5)+
  labs(x = "Weeks at birth", y = "Birthweight (pounds)")+
  scale_color_manual(values = c("purple", "orange"), name = "Smoking Status")+
  scale_size_continuous(name = "Weight Gain")


bw_scatter
```

####Histograrms and qq-plots

Create side-by-side histograms (hint: facet_wrap) for observations of baby weight pregnancy. geom_histogram: one variable (x = )

```{r}
# creating exploratory histogram of birthweights
bw_hist <- ggplot(nc_new, aes(x = tpounds))+
  geom_histogram(aes(fill=sdp))+
  facet_wrap(~ sdp, scale = "free")

bw_hist

# craete exploratory qq plots
bw_qq <- ggplot(nc_new, aes(sample=tpounds))+
  geom_qq()+
  facet_wrap(~sdp, scale = "free")

bw_qq

# Summary table of mean, sd, max, min

birth_summary <- nc_new %>% 
  group_by(sdp) %>% 
  summarize(
    mean_wt = mean(tpounds),
    sd_wt = sd(tpounds),
    max_wt = max(tpounds)
  )

birth_summary
```

###4. Two-sample t-tests

Use the pull() function to extract column information as a VECTOR
```{r}
# create a subset from nc_new data that only contains observations for babies born to smoking mothers, and then pulls JUST the "tpounds" column as a vector

s <- nc_new %>% 
  filter(sdp == "smoker") %>% 
  pull(tpounds)

# creating a vector containing only birthwieghts for babies born to nonsmoking mothers "ns".

ns <- nc_new %>% 
  filter(sdp == "nonsmoker") %>% 
  pull(tpounds)
```

Is there a significant difference in birthweights for babies born to smoking and nonsmoking mothers?

```{r}

# H0: Means are equal (difference in means = 0)
# HA: Means are NOT equal (difference in means is NOT 0)

# First, let's check for equal variance
# F: ratio of sample variances
# H0: Ratio of variance = 1
# HA: Ratio of variance is NOT = 1

# performing an F-test for equal variances
f_test <- var.test(s, ns)
f_test # p = 0.197, variances are equal


```

# we are going to override the default var.equal = TRUE in the 


```{r}
t_diff <- t.test(s, ns, var.equal = TRUE)
t_diff

# If I ask "Are babies born to smoking mothers significantly smaller (in birthweight) than those born to nonsmoking mothers?"

t_less <- t.test(s, ns, var.equal = TRUE, alternative = "less")
t_less 

# t_more <- t.test(ns, s, var.equal = TRUE, alternative = "greater")

```

LaTex
Inline referencing examples:

(t(df) = ??, p = ??, alpha = 0.05)

... weights differed significantly (t(`r{t_diff$parameter}`)=`r{round(t_diff$statistic,2)}`, *p* = `r {t_diff$p.value}`, $\alpha$)

###5. Effect size (Cohen's *d*)

```{r}
effect_size <- cohen.d(ns,s)
effect_size
```


